SU($n$) ferromagnets near the Heisenberg point

### 2020-06-14 collective_sun/sun_perturbation.tex
added high-level description of the construction of excited states in the general SU(n) case

\title{SU($n$) ferromagnets near the Heisenberg critical point}%
\author{Michael A. Perlin}%
\date{\today}

@@

As a concrete example, this sort of scenario can be used in the SU(2)
case to generate spin squeezed states that harness entanglement to
surpass classical limits on measurement
precision\cite{he2019engineering, perlin2020spin}.
Starting with a general (gapped) SU(2)-symmetric Hamiltonian of the
form in \eqref{eq:H_SU2} we can consider the addition of Ising
interactions
\begin{align}
  H_{\t{Ising}} = \sum_{i\ne j} J_{ij} s_\z^{(i)} s_\z^{(j)}.
  \label{eq:Ising}
\end{align}
If these interactions are sufficiently weak for perturbative
treatment, as in the case of the XXZ model near the Heisenberg
critical point, then the effective Hamiltonian induced by
$H_{\t{Ising}}$ on initially PS states $\ket\psi\in\M_0$ is the
one-axis twisting (OAT) Hamiltonian\cite{ma2011quantum}
\begin{align}
  H_{\t{OAT}} = \chi S_\z^2,
  \label{eq:OAT}
\end{align}
where the collective operator $S_\z\equiv\sum_{j=1}^Ns_\z^{(j)}$, and
the squeezing strength $\chi$ is equal to the mean Ising coupling
strength $J_{ij}$.  If Ising interactions are unavailable, the same
effect can be achieved by applying an inhomogeneous magnetic field,
\begin{align}
  H_{\t{B}} = \sum_i B_i s_\z^{(i)},
  \label{eq:magnetic}
\end{align}
which gives rise to an OAT Hamiltonian at second order in perturbation
theory, with a squeezing strength $\chi$ that is related to the
variance of the magnetic field strength $B_i$.

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta>0$ with respect to $H_0$.  Rather
than treating these excited states as ``intermediate'' or ``virtual''
in the derivation of an effective ground-state Hamiltonian, we can
perform numerical simulations in the joint Hilbert space of both
ground and excited states.  When the perturbation $\O\p{\m w,X}$ is
small, namely with operator norm
$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
equivalent (up to perturbative corrections) to a higher-order
perturbation theory.  Moreover, the first-class treatment of excited
states in finite systems enables simulating the dynamics induced by
{\it non-perturbative} additions $\O\p{\m w,X}$ to the base
Hamiltonian $H_0$.

In order to simulate dynamics within excited-state manifolds, we first
need to classify the set of excited states in order to construct a
suitable basis.  In the case of SU(2), a common starting point is to
identify states with a definite (non-negative) spin length
$S=\set{N/2,N/2-1,\cdots}$, and a spin projection
$m_\z\in\set{S,S-1,\cdots,-S}$ onto a quantization axis.  To
generalize this classification to SU($n$), we denote the manifold of
states with definite fixed spin length $S$ by $\M_{N/2-S}$.  The PS
manifold of Dicke states with $S=N/2$ is then $\M_0$, the manifold of
single-magnon spin-wave states with $S=N/2-1$ is $\M_1$, and more
generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Rather than the spin projection $m_\z$ we then identify states by
labels $m=\p{m_\up,m_\dn}\in\A_2\p{N}$ in which $m_\mu$ denotes the
occupation number of the single-spin state $\mu$, with
$m_\up-m_\dn=m_\z$ and $m_\up+m_\dn=N$.

The label $m$ suffices to identify the PS (Dicke) manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional
labels to uniquely identify states in the $r$-magnon manifold $\M_r$
with $r>0$.  Single-magnon ($r=1$) spin-wave states of SU(2) spin
systems can be written in the form
\begin{align}
  \ket{m,\m v} \equiv \O\p{\m v,\sigma_\z} \ket{m}
  = \sum_j v_j \sigma_\z^{(j)} \ket{m},
\end{align}
where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
If an SU(2) spin system is translationally invariant, then these
spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}/\sqrt{N}$ with
wavenumber $k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2)
take the form
$\ket{m,\m v}\equiv\O\p{\m v,\sigma_\z^{\otimes r}}\ket{m}$, where now
$\m v$ is an $r$-index tensor that encodes the structure of
correlations in the excited state $\ket{m,\m v}$ (see Figure
\ref{fig:structure}).  Note that $M$-body operators will generally
couple the PS manifold $\M_0$ to all $r$-magnon manifolds $\M_r$ with
$r\le M$ (see Figure \ref{fig:manifold_sketch}).  To generalize the
notion of $r$-magnon manifolds $\M_r$ to the case of SU($n$), where
there is no obvious notion of a spin length $S$, we therefore define
$\M_r$ for $r>0$ as the manifold of states that (i) can be generated
by the action of $r$-body operators on $\M_0$, and (ii) are orthogonal
to all states in $\bigoplus_{s<r}\M_s$.

\begin{figure}
  \centering
  \includegraphics{structure/mat_L101_a2.pdf}
  \caption{Visualizations of dimension-$r$ tensors $\m v$ that
    generate the lowest-lying $r$-magnon states
    $\ket{\m v,m} \equiv \O\p{\m v, \sigma_\z^{\otimes{r}}} \ket{m}$
    with definite excitation energy $\Delta$ above the PS manifold
    $\M_0$ of a periodic 1D lattice of $N=101$ spins interacting via
    the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
    couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
    minimal distance between spins $i,j$ on a periodic 1D lattice.
    Top row ($r=2$) shows the full 2D matrix
    $\m v = \sum_{i,j}v_{ij}\op{i}{j}$, and bottom row ($r=3$) shows a
    single 2D slice of the full 3D tensor,
    $\tilde{\m v}_0 = \sum_{i,j}v_{ij,0}\op{i}{j}$.  Blue (red) pixels
    correspond to matrix elements with a positive (negative) sign, and
    darker colors indicate larger absolute values.  In the case of
    $r=2$, translational invariance enforces $v_{ij}=v_{i-d,j-d}$ for
    all displacements $d$, so every row of $\m v$ is some translation
    of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
    of the full 3D tensor can be obtained by appropriately translating
    the 2D slice $\tilde{\m v}_0$ shown above.  In the case of SU($n$)
    with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
    tensors $\m v$ remain identical, but now correspond to a
    degenerate set of states that are generated by different choices
    of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
  \label{fig:structure}
\end{figure}

\begin{figure}
  \centering
  \begin{tikzpicture}
    \draw[thickest] (0,0) -- ++(10em,0);
    \foreach \hh in {0.15em,0.05em,-0.05em,-0.15em} {
      \draw (0,3em+\hh) -- ++(10em,0);
    }
    \draw[->] (7em,.25em) arc (220:0:.4em);
    \draw[->] (7em,.25em) to [bend left]
    node[left] {$\O\p{\m v,\sigma_\z}$} (7em,2.7em);
    \node () at (5em,4.7em) {$\vdots$};
    \foreach \hh in {0.25em,0.15em,0.05em,-0.05em,-0.15em,-0.25em} {
      \draw (0,6em+\hh) -- ++(10em,0);
    }
    \draw[->] (-.4em,0em) arc (220:0:.4em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,3em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,4.3em);
    \draw[->] (-.4em,0) to [bend left, in=130]
    node[left] {$\O\p{\m w,\sigma_\z^{\otimes M}}$} (-.2em,6em);
    \node[right] () at (10em,0) {$\M_0,S=\frac{N}{2}$};
    \node[right] () at (10em,3em) {$\M_1,S=\frac{N}{2}-1$};
    \node[right] () at (10em,6em) {$\M_M,S=\frac{N}{2}-M$};
  \end{tikzpicture}
  \caption{Schematic of relationships between $r$-magnon manifolds
    $\M_r$ with total spin $S=N/2-r$ in an SU(2) spin system.
    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the
    ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$
    as well as $\M_0$ itself, while $M$-body operators
    $\O\p{\m w,\sigma_\z^{\otimes M}}$ couple $\M_0$ to all $r$-magnon
    manifolds $\M_r$ with $r\le M$.}
  \label{fig:manifold_sketch}
\end{figure}

\red{[note: this section is not yet finished]}

@@

\red{[look at \cite{rai1988correlated} for classifying asymmetric
  states]}

@@

@@

\red{[todo: make this a subsection of Appendix
  \ref{sec:single_body_eigenstates}]}

Here we show that in $D$ spatial dimensions, power-law
SU($n$)-symmetric interactions of the form in \eqref{eq:H_0} with
$h_{ij}=-1/\abs{i-j}^\alpha$ yield a non-vanishing spectral gap when
$\alpha\le D$.  For definiteness, we will consider a periodic
hypercubic lattice of $N=L^D$ spins.  The spectral gap of the
interaction Hamiltonian is then (see Appendix
\ref{sec:single_body_eigenstates})
\begin{align}
  \Delta
  = \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
  = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
\end{align}
where $\ZZ_L$ denotes the set of integers modulo $L$;
$k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a single-magnon
spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic
lattice is implicitly understood to mean the smallest Euclidean
distance of $d$ from a fixed origin.  In order to yield the smallest
excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should
maximize the contribution of the cosine term above, which is achieved
by a wavenumber that minimizes the oscillations of this term when
integrated over the entire lattice.  A suitable candidate for a
minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which
corresponds to an excitation energy
\begin{align}
  \Delta = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
\end{align}
Defining $\epsilon\equiv2/L$ and a rescaled domain symmetrized about
$0$, $\SS_\epsilon\simeq\ZZ_L/\epsilon$ with
$\SS_\epsilon\subset\sp{-1,1}$, we substitute $x\simeq\epsilon d$ to
get
\begin{align}
  \Delta
  = \sum_{\substack{x\in\SS_\epsilon^D\\\abs{x}\ge\epsilon}}
  \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
  = \epsilon^{\alpha-D} \sum_{\substack{x\in\SS_L^D\\\abs{x}\ge\epsilon}}
  \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
\end{align}
As $\epsilon\to0$ the discrete sum over $x$ is well approximated by an
integral that avoids an infinitesimal region about the origin, i.e.
\begin{align}
  \Delta = \epsilon^{\alpha-D} \I_D\p{\epsilon},
  &&
  \I_D\p{\epsilon}
  \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
  \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha},
\end{align}
where the integral $\I_D\p{\epsilon}$ is defined using the symmetric
interval $\TT_a\equiv\p{-a/2,a/2}$.  The integrand of
$\I_D\p{\epsilon}$ is strictly positive and well-behaved on the
entirety of its domain except for the origin, where depending on the
value of $\alpha$ the integrand may vanish or diverge as
$\abs{x}\to0$.  Together, these facts mean that
\begin{align}
  \I_D\p{\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
  &&
  \Delta \stackrel{\epsilon\to0}{\sim} \epsilon^{\alpha-D-\gamma},
\end{align}
for some $\gamma\ge0$, which implies that gap $\Delta$ is
non-vanishing when $\alpha\le D\le D+\gamma$.

@@

@@

A diagram $g\in\G\p{\m M}$ with $\m M=\p{M_1,M_2,\cdots,M_p}$ consists
of $p$ circles, and $g_S$ dots at the intersection of circles
$\alpha\in S\subset\ZZ_p$.  By slight abuse of notation, labeling each
circle $\alpha$ by a tensor $w_\alpha$ defines the scalar
$g\p{\m w}\equiv\ut{\m w}\p{g}$.  If, for example, all tensors
$w_\alpha$ are all permutationally symmetric, then
\begin{align}
  \diagram{example_w123}
  = \sum_{\p{a,b,c,d,e}\in\D_N\p{5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
  \label{eq:example_123}
\end{align}
When the tensors $\m w$ are not all permutationally symmetric, one can
additionally equip a diagram $g$ with a specific index assignment
$\ell\in\L\p{g}$, in which case the diagram is equal to
$\ell\p{\m w}\equiv\col{\m w}\p{\ell}$.  Diagrams that are not
equipped with a specific index assignment $\ell$ are then an average
over all $\ell\in\L\p{g}$,
i.e.~$g\p{\m w}\equiv\sum_{\ell\in\L\p{g}} \ell\p{\m w} /
\abs{\L\p{g}}$.  To keep mathematical expressions simple, we will use
symmetric tensors in all examples in this section, but our discussions
apply just as well to the asymmetric case.

>>>>

\title{SU($n$) ferromagnets near the Heisenberg point}
\author{Michael A. Perlin}
\date{\today}

@@

As a concrete example, this sort of scenario can be used in the SU(2)
case to generate spin squeezed states that harness entanglement to
surpass classical limits on measurement
precision\cite{he2019engineering, perlin2020spin}.
Starting with a general (gapped) SU(2)-symmetric Hamiltonian of the
form in \eqref{eq:H_SU2} we can consider the addition of Ising
interactions
\begin{align}
  H_{\t{Ising}} = \sum_{i\ne j} J_{ij} s_\z^{(i)} s_\z^{(j)}.
  \label{eq:Ising}
\end{align}
If these interactions are sufficiently weak for perturbative
treatment, as in the case of the XXZ model near the Heisenberg point,
then the effective Hamiltonian induced by $H_{\t{Ising}}$ on initially
PS states $\ket\psi\in\M_0$ is the one-axis twisting (OAT)
Hamiltonian\cite{ma2011quantum}
\begin{align}
  H_{\t{OAT}} = \chi S_\z^2,
  \label{eq:OAT}
\end{align}
where the collective operator $S_\z\equiv\sum_{j=1}^Ns_\z^{(j)}$, and
the squeezing strength $\chi$ is equal to the mean Ising coupling
strength $J_{ij}$.  If Ising interactions are unavailable, the same
effect can be achieved by applying an inhomogeneous magnetic field,
\begin{align}
  H_{\t{B}} = \sum_i B_i s_\z^{(i)},
  \label{eq:magnetic}
\end{align}
which gives rise to an OAT Hamiltonian at second order in perturbation
theory, with a squeezing strength $\chi$ that is related to the
variance of the magnetic field strength $B_i$.

@@

In order to find the effective Hamiltonian $H_{\t{eff}}^{(2)}$ induced
by the $M$-body operator $\O\p{\m w,X}$ on the PS manifold $\M_0$ at
second order in perturbation theory, we constructed states with
definite excitation energy $\Delta>0$ with respect to $H_0$.  Rather
than treating these excited states as ``intermediate'' or ``virtual''
in the derivation of an effective ground-state Hamiltonian, we can
perform numerical simulations in the joint Hilbert space of both
ground and excited states.  When the perturbation $\O\p{\m w,X}$ is
small, namely with operator norm
$\norm{\O\p{\m w,X}}<\Delta_{\t{gap}}/2$, such simulations are
equivalent (up to vanishing perturbative corrections) to a
higher-order perturbation theory.  Moreover, the first-class treatment
of excited states in finite systems enables simulating the dynamics
induced by {\it non-perturbative} additions $\O\p{\m w,X}$ to the base
Hamiltonian $H_0$.

In order to simulate dynamics within excited-state manifolds, we first
need to classify the set of excited states, and construct a suitable
basis.  In the case of SU(2), a common starting point is to identify
states with a definite (non-negative) spin length
$S=\set{N/2,N/2-1,\cdots}$, and a spin projection
$m_\z\in\set{S,S-1,\cdots,-S}$ onto a quantization axis.  To
generalize this classification to SU($n$), we denote the manifold of
states with definite fixed spin length $S$ by $\M_{N/2-S}$.  The PS
manifold of Dicke states with $S=N/2$ is then $\M_0$, the manifold of
single-magnon spin-wave states with $S=N/2-1$ is $\M_1$, and more
generally the manifold of $r$-magnon states with $S=N/2-r$ is $\M_r$.
Rather than the spin projection $m_\z$ we then identify states by
labels $m=\p{m_\up,m_\dn}\in\A_2\p{N}$ in which $m_\mu$ denotes the
occupation number of the single-spin state $\mu$, with
$m_\up+m_\dn=N$ and $m_\up-m_\dn=m_\z$.

\begin{figure}
  \centering
  \includegraphics{structure/mat_L101_a2.pdf}
  \caption{Visualizations of dimension-$r$ tensors $\m w$ that
    generate the lowest-lying (unnormalized) $r$-magnon states
    $\ket{\m w,m} \equiv \O\p{\m w, \sigma_\z^{\otimes{r}}} \ket{m}$
    with definite excitation energy $\Delta$ above the PS manifold
    $\M_0$ of a periodic 1D lattice of $N=101$ spins interacting via
    the SU(2)-symmetric Hamiltonian in \eqref{eq:H_SU2} with power-law
    couplings $h_{ij}=-1/\oper{dist}\p{i,j}^2$.  Here
    $\oper{dist}\p{i,j}\equiv\min\p{\abs{i-j},N-\abs{i-j}}$ is the
    minimal distance between spins $i,j$.  Top row ($r=2$) shows the
    full 2D matrix $\m w = \sum_{i,j}w_{ij}\op{i}{j}$, and bottom row
    ($r=3$) shows a single 2D slice of the full 3D tensor,
    $\tilde{\m w}_0 = \sum_{i,j}w_{ij,0}\op{i}{j}$.  Blue (red) pixels
    correspond to matrix elements with a positive (negative) sign, and
    darker colors indicate larger absolute values.  In the case of
    $r=2$, translational invariance enforces $w_{ij}=w_{i-d,j-d}$ for
    all displacements $d$, so every row of $\m w$ is some translation
    of a single (arbitrary) row.  Similarly, when $r=3$ every 2D slice
    of the full 3D tensor can be obtained by appropriately translating
    the 2D slice $\tilde{\m w}_0$ shown above.  In the case of SU($n$)
    with $n>2$, the ``$r$-magnon'' excitation energies $\Delta$ and
    tensors $\m w$ remain identical, but now correspond to a
    degenerate set of states that are generated by different choices
    of $r$-body operators in place of $\sigma_\z^{\otimes r}$.}
  \label{fig:structure}
\end{figure}

The label $m$ suffices to identify the PS (Dicke) manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_2\p{N}}$, but we need additional
labels to uniquely identify states in the $r$-magnon manifold $\M_r$
with $r>0$.  Single-magnon ($r=1$) spin-wave states of SU(2) spin
systems can be written in the form
\begin{align}
  \ket{m,\m v} \equiv \O\p{\m v,\sigma_\z} \ket{m}
  = \sum_j v_j \sigma_\z^{(j)} \ket{m},
\end{align}
where each $v_j$ is a scalar and $\sigma_\z$ is a Pauli-$z$ operator.
If an SU(2) spin system is translationally invariant, then these
spin-wave eigenstates have $v_j=e^{\i k_{\m v}\c j}$ with wavenumber
$k_{\m v}\ne0$.  Similarly, $r$-magnon states for SU(2) take the form
$\ket{m,\m w}\equiv\O\p{\m w,\sigma_\z^{\otimes r}}\ket{m}$, where now
$\m w$ is an $r$-index tensor that encodes the structure of
correlations in the excited state $\ket{m,\m w}$ (see Figure
\ref{fig:structure}).  Considering states of this form, built using
tensor products of Pauli-$z$ ($\sigma_\z$) operators, suffices to
construct a basis for all excited states of an SU(2) spin system (see
Appendix \ref{sec:operator_content}).  Note that $r$-body operators
with arbitrary coefficients $\m w$ will generally couple the PS
manifold $\M_0$ to all $q$-magnon manifolds $\M_q$ with $q\le r$ (see
Figure \ref{fig:manifold_sketch}).  In order to generalize the notion
of $r$-magnon manifolds $\M_r$ to the case of SU($n$) with $n>2$,
where the analogous notion of a spin length $S$ is less intuitive, we
therefore define $\M_r$ for $r>0$ as the manifold of states that (i)
can be generated by the action of $r$-body operators on $\M_0$, and
(ii) are orthogonal to all states in $\bigoplus_{q<r}\M_q$.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \draw[thickest] (0,0) -- ++(10em,0);
    \foreach \hh in {0.15em,0.05em,-0.05em,-0.15em} {
      \draw (0,3em+\hh) -- ++(10em,0);
    }
    \draw[->] (7em,.25em) arc (220:0:.4em);
    \draw[->] (7em,.25em) to [bend left]
    node[left] {$\O\p{\m v,\sigma_\z}$} (7em,2.7em);
    \node () at (5em,4.7em) {$\vdots$};
    \foreach \hh in {0.25em,0.15em,0.05em,-0.05em,-0.15em,-0.25em} {
      \draw (0,6em+\hh) -- ++(10em,0);
    }
    \draw[->] (-.4em,0) arc (220:0:.4em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,3em);
    \draw[->] (-.4em,0) to [bend left, in=130] (-.2em,4.3em);
    \draw[->] (-.4em,0) to [bend left, in=130]
    node[left] {$\O\p{\m w,\sigma_\z^{\otimes r}}$} (-.2em,6em);
    \node[right] () at (10em,0) {$\M_0,S=\frac{N}{2}$};
    \node[right] () at (10em,3em) {$\M_1,S=\frac{N}{2}-1$};
    \node[right] () at (10em,6em) {$\M_r,S=\frac{N}{2}-r$};
  \end{tikzpicture}
  \caption{Schematic of relationships between $q$-magnon manifolds
    $\M_q$ with total spin $S=N/2-q$ in an SU(2) spin system.
    Single-body operators $\O\p{\m v,\sigma_\z}$ generally couple the
    ground-state manifold $\M_0$ to the single-magnon manifold $\M_1$
    as well as $\M_0$ itself, while $r$-body operators
    $\O\p{\m w,\sigma_\z^{\otimes r}}$ couple $\M_0$ to all $q$-magnon
    manifolds $\M_q$ with $q\le r$.}
  \label{fig:manifold_sketch}
\end{figure}

In the case of an SU($n$) spin system, we can similarly start with the
ground-state PS manifold
$\M_0=\oper{span}\set{\ket{m}:m\in\A_n\p{N}}$, and generate excited
states by the action of multi-body operators, $\O\p{\m w,X}\ket{m}$.
The choice of tensors $\m w$ for generating states in the $r$-magnon
manifold $\M_r$ is the same as before, but we now have the additional
complication of having to choose $r$-spin operators, $X$.  As with the
choice of $\sigma_\z^{\otimes r}$ for SU(2), it suffices to choose
tensor products of {\it diagonal} single-spin operators (that is,
diagonal in a fixed single-particle basis), but there will now be
$\sim\p{n-1}^r$ distinct choices that generate degenerate, orthogonal
states: one of $n-1$ diagonal operators (that are linearly independent
of each other and the identity) for each of $r$ tensor factors.  We
discuss the exact choice of operators in more detail in Appendix
\ref{sec:operator_content}.  To summarize, for each PS state
$\ket{m}$, we identify a basis $\set{Z_{m\nu}:\nu\in\ZZ_{n-1}}$ of
single-spin operators that are diagonal and linearly independent of
the identity, and choose a list of integers
$z\equiv\p{z_1,z_2,\cdots,z_{n-1}}\in\A_{n-1}\p{r}$ to define the
$r$-spin operator
\begin{align}
  X_{mz} \equiv \bigotimes_{\nu=1}^{n-1} Z_{m\nu}^{\otimes z_\nu}
  = \underbrace{Z_{m,1} \otimes \cdots \otimes Z_{m,1}}_{z_1~\t{times}}
  \otimes
  \underbrace{Z_{m,2} \otimes \cdots \otimes Z_{m,2}}_{z_2~\t{times}}
  \otimes \cdots \otimes
  \underbrace{Z_{m,n-1} \otimes \cdots
    \otimes Z_{m,n-1}}_{z_{n-1}~\t{times}}.
\end{align}
Our basis for $\M_r$ then consists of the states
$\ket{m,\m w,z}\equiv\O\p{\m w,X_{mz}}\ket{m}$.

\red{[todo: discuss the use of this basis in numerical simulations]}

@@

\red{[look at \cite{rai1988correlated} for classifying asymmetric
  states?]}

@@

\begin{acknowledgments}
  We thank Asier Pi\~neiro Orioli for helpful discussions and
  technical contributions to this work.
\end{acknowledgments}

@@

Here we show that in $D$ spatial dimensions, power-law
SU($n$)-symmetric interactions of the form in \eqref{eq:H_0} with
$h_{ij}=-1/\abs{i-j}^\alpha$ yield a non-vanishing spectral gap when
$\alpha\le D$.  For definiteness, we will consider a periodic
hypercubic lattice of $N=L^D$ spins.  The spectral gap of the
interaction Hamiltonian is then determined by well-known spin-wave
theory (see Ref.~[\citenum{ashcroft1978solid}] or Appendix
\ref{sec:single_body_eigenstates})
\begin{align}
  \Delta
  = \sum_{d\in\ZZ_L^D} h_{0,d} \sp{\cos\p{d \c k_{\t{SE}}}-1}
  = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d \c k_{\t{SE}}}}{\abs{d}^\alpha},
\end{align}
where $\ZZ_L$ denotes the set of integers modulo $L$;
$k_{\t{SE}}\in\ZZ_L^D\times2\pi/L$ is a wavenumber for a single-magnon
spin-wave state; and the norm $\abs{d}$ for a vector $d$ on a periodic
lattice is implicitly understood to mean the smallest Euclidean
distance of $d$ from a fixed origin.  In order to yield the smallest
excitation energy $\Delta$, the wavenumber $k_{\t{SE}}$ should
maximize the contribution of the cosine term above, which is achieved
by a wavenumber that minimizes the oscillations of this term when
integrated over the entire lattice.  A suitable candidate for a
minimal wavenumber is $k_{\t{SE}}=\p{2\pi/L,0,0,\cdots}$, which
corresponds to an excitation energy
\begin{align}
  \Delta = \sum_{\substack{d\in\ZZ_L^D\\\abs{d}\ge1}}
  \f{1-\cos\p{d_12\pi/L}}{\abs{d}^\alpha}.
\end{align}
Defining $\epsilon\equiv2/L$ and a rescaled domain symmetrized about
$0$, $\SS_\epsilon\simeq\ZZ_L/\epsilon$ with
$\SS_\epsilon\subset\sp{-1,1}$, we substitute $x\simeq\epsilon d$ to
get
\begin{align}
  \Delta
  = \sum_{\substack{x\in\SS_\epsilon^D\\\abs{x}\ge\epsilon}}
  \f{1-\cos\p{\pi x_1}}{\abs{x/\epsilon}^\alpha}
  = \epsilon^{\alpha-D} \sum_{\substack{x\in\SS_L^D\\\abs{x}\ge\epsilon}}
  \epsilon^D \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha}.
\end{align}
As $\epsilon\to0$ the discrete sum over $x$ is well approximated by an
integral that avoids an infinitesimal region about the origin, i.e.
\begin{align}
  \Delta = \epsilon^{\alpha-D} \I_D\p{\epsilon},
  &&
  \I_D\p{\epsilon}
  \equiv \int_{\TT_1^D\setminus\TT_\epsilon^D} \d^Dx\,
  \f{1-\cos\p{\pi x_1}}{\abs{x}^\alpha},
\end{align}
where the integral $\I_D\p{\epsilon}$ is defined using the symmetric
interval $\TT_a\equiv\p{-a/2,a/2}$.  The integrand of
$\I_D\p{\epsilon}$ is strictly positive and well-behaved on the
entirety of its domain except for the origin, where depending on the
value of $\alpha$ the integrand may vanish or diverge as
$\abs{x}\to0$.  Together, these facts mean that
\begin{align}
  \I_D\p{\epsilon} \stackrel{\epsilon\to0}{\sim} \epsilon^{-\gamma},
  &&
  \Delta \stackrel{\epsilon\to0}{\sim} \epsilon^{\alpha-D-\gamma},
\end{align}
for some $\gamma\ge0$, which implies that gap $\Delta$ is
non-vanishing when $\alpha\le D\le D+\gamma$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Choosing the operator content}
\label{sec:operator_content}

\red{[todo: write section]}

@@

A diagram $g\in\G\p{\m M}$ with $\m M=\p{M_1,M_2,\cdots,M_p}$ consists
of $p$ circles, and $g_S$ dots at the intersection of circles
$\alpha\in S\subset\ZZ_p$.  By slight abuse of notation, labeling each
circle $\alpha$ by a tensor $w_\alpha$ defines the scalar
$g\p{\m w}\equiv\ut{\m w}\p{g}$.  If, for example, all tensors
$w_\alpha$ are all permutationally symmetric, then
\begin{align}
  \diagram{example_w123}
  = \sum_{\p{a,b,c,d,e}\in\D_N\p{5}}
  w_1\p{a,b,c} w_2\p{b,d,e} w_3\p{b,c,d,e}.
  \label{eq:example_123}
\end{align}
When the tensors $\m w$ are not all permutationally symmetric, one can
additionally equip a diagram $g$ with a specific index assignment
$\ell\in\L\p{g}$, in which case the diagram is equal to
$\ell\p{\m w}\equiv\col{\m w}\p{\ell}$.  Diagrams that are not
equipped with a specific index assignment $\ell$ are then an average
over all $\ell\in\L\p{g}$,
i.e.~$g\p{\m w}\equiv\sum_{\ell\in\L\p{g}} \ell\p{\m w} /
\abs{\L\p{g}}$.  To keep mathematical expressions simple, we will use
symmetric tensors in all examples throughout this section, but our
discussions apply just as well to the asymmetric case.
